Bayesian Sparsification Methods for Deep Complex-valued Networks

### 2020-06-23 paper/paper.tex
12 cut off some text and shrunk acknowledgements piece

The paper is structured as follows. Sec.~\ref{sec:variational_dropout} reviews Variational
Dropout, and sec.~\ref{sec:c_valued_networks} provides a brief summary of the inner
workings of complex-valued networks. The main contribution of this study is presented
in sec.~\ref{sec:c_variational_dropout}, where we provide the details of $\cplx$-valued
variational sparsification method. In sec.~\ref{sec:experiments} we estimate the compression
and performance trade-off on shallow and deep $\cplx$-valued networks, and discuss the
outcomes.

@@

With a special posterior approximation family $q_\theta$ and suitable prior $\pi$,
Variational Inference can be used as model regularization and sparsification technique.

@@

During the \emph{``sparsify''} stage we make every layer stochastic and apply $\cplx$-valued
variational sparsification (sec.~\ref{ssub:vd_prior}, ~\ref{ssub:ard_prior}) or its $\real$
counterpart. We inject a coefficient $
  C \in (0, 1]
$ at the KL divergence term in \eqref{eq:sgvb_estimator}:
\begin{equation}  \label{eq:elbo_with_coef}
  - \frac{C}N KL(q_{\theta} \| \pi_{\lambda})
  + \frac1{M} \sum_{k=1}^M
      \log p_{\phi}(x_{i_k} \mid g(\varepsilon_k; \theta))
  \,.
  \tag{\ref{eq:sgvb_estimator}'}
\end{equation}
In contrast to \cite{molchanov_variational_2017}, who anneal $C$ from zero to one during
training, we use constant $C$ and vary it between runs. This allows us to explore the
compression-performance profile by balancing model's likelihood and posterior's penalty
for diverging form the sparsifying prior in \eqref{eq:elbo_with_coef}. In particular,
higher $C$ implies higher sparsity.
% RFC: citation needed fot this "trick". Cite $\beta$-VAE maybe?

@@

Having verified the $\cplx$ variational sparsification method on MNIST-like datasets
and simple models, we turn to the CIFAR10 dataset comprising $32\times 32$ colour
images of $10$ classes \citep{krizhevsky_learning_2009} and focus on the VGG16 network
\citep{simonyan_very_2015}.
%
% same experimental setup as in MNIST, describe the VGG16 model and its cplx-valued modification
We train the VGG16 network and its $\cplx$ variant, in which we have replaced $\real$-valued
layers with their $\cplx$-valued counterparts. We do not halve or double the features
in any network, since the goal of this experiment is to assess the trade-off for a deep
convolutional network.
%
% differences with mnist
Unlike experiment in sec.~\ref{sub:mnist_like_datasets}, we consider the raw features
only, use full training split and allocate $20$, $40$, and $20$ epochs to each stage.
During training every mini-batch of $128$ samples is augmented by random horizontal flipping
and random cropping, which is done by zero-padding the image with four pixels and extracting
a $32\times 32$ patch from the $40\times 40$ intermediate image.
%
We vary $C$ in \eqref{eq:elbo_with_coef} over $
  \{\tfrac32 2^{-\tfrac{k}2} \colon k=7, \cdots, 15\}
$, compare $\real / \cplx$ VD and ARD methods and measure accuracy on the usual test split.

@@

% draw some conclusions
The compression-accuracy curve in figure~\ref{fig:figure__cifar10__trade-off} shows that
it is possible to confidently achieve around $\times100$ compression of a deep $\cplx$VNN
without losing accuracy, provided the network is fine-tuned after undergoing Variational
sparsification. Regarding methods themselves, $\cplx$-VD and $\cplx$-ARD follow the same
declining compression-accuracy pattern, but for the same setting $C$ the latter provides
slightly less compression with marginally better accuracy.

@@

The results of this study have direct implications for embedded deep learning applications
both in terms of lower storage requirements and higher throughput stemming from fewer
floating point multiplications due to sparsity, despite somewhat higher arithmetic
complexity of $\cplx$-valued networks.

@@

% the final camera-ready version can (and probably should) include acknowledgements
% Typically, this will include thanks to reviewers who gave useful comments, to colleagues
% who contributed to the ideas, and to funding agencies and corporate sponsors that provided
% financial support.
We would like to thank the anonymous reviewers for their useful comments, and Evgenii
Egorov, Ruslan Kostoev (ADASE) and Danila Doroshin (Huawei) for fruitfull discussions.
% We thank Skoltech CDISE HPC Zhores cluster staff for providing computational infrastructure.
The authors acknowledge the use of the Skoltech CDISE HPC cluster ``Zhores'' for obtaining
the results presented in this paper.

>>>>

The paper is structured as follows. Sec.~\ref{sec:variational_dropout} reviews Variational
Dropout, and sec.~\ref{sec:c_valued_networks} provides a brief summary of the inner
workings of complex-valued networks. The main contribution of this study is presented
in sec.~\ref{sec:c_variational_dropout}, where we provide the details of $\cplx$-valued
variational sparsification methods. In sec.~\ref{sec:experiments} we estimate the compression
and performance trade-off on shallow and deep $\cplx$-valued networks, and discuss the
outcomes.

@@

Variational Inference can be used as model regularization and sparsification method
for certain posterior approximation $q_\theta$ and prior $\pi$.

@@

During the \emph{``sparsify''} stage we make every layer stochastic and apply variational
sparsification (sec.~\ref{ssub:vd_prior}, ~\ref{ssub:ard_prior}, or their $\real$ versions).
We inject a coefficient $
  C \in (0, 1]
$ at the KL divergence term in \eqref{eq:sgvb_estimator}:
\begin{equation}  \label{eq:elbo_with_coef}
  - \frac{C}N KL(q_{\theta} \| \pi_{\lambda})
  + \frac1{M} \sum_{k=1}^M
      \log p_{\phi}(x_{i_k} \mid g(\varepsilon_k; \theta))
  \,.
  \tag{\ref{eq:sgvb_estimator}'}
\end{equation}
In contrast to \cite{molchanov_variational_2017}, who anneal $C$ from zero to one during
training, we use constant $C$ and vary it between runs. This allows us to explore the
compression-performance profile by balancing model's likelihood and posterior's penalty
for diverging form the sparsifying prior in \eqref{eq:elbo_with_coef}. In particular,
higher $C$ implies higher sparsity.
% RFC: citation needed fot this "trick". Cite $\beta$-VAE maybe?

@@

Having verified the $\cplx$ variational sparsification method on MNIST-like datasets
and simple models, we turn to the CIFAR10 dataset comprising $32\times 32$ colour
images of $10$ classes \citep{krizhevsky_learning_2009} and focus on the VGG16 network
\citep{simonyan_very_2015}.
%
% same experimental setup as in MNIST, describe the VGG16 model and its cplx-valued modification
We train the VGG16 network and its $\cplx$ variant, in which we have replaced $\real$-valued
layers with their $\cplx$-valued counterparts. We do not halve or double the features
in any network, since the goal of this experiment is to assess the trade-off for a deep
convolutional network.
%
% differences with mnist
Unlike experiment in sec.~\ref{sub:mnist_like_datasets}, we consider the raw features
only, use full training split, measure accuracy on the usual test split, and allocate
$20$, $40$, and $20$ epochs to each stage.
During training every mini-batch of $128$ samples is augmented by random horizontal flipping
and random cropping, which is done by zero-padding the image with four pixels and extracting
a $32\times 32$ patch from the $40\times 40$ intermediate image.

@@

% draw some conclusions
The compression-accuracy curve in figure~\ref{fig:figure__cifar10__trade-off}, constructed
for $C = \tfrac32 2^{-\tfrac{k}2}$ with $k=7, \cdots, 15$, shows that
it is possible to confidently achieve around $\times100$ compression of a deep $\cplx$VNN
without losing accuracy, provided the network is fine-tuned after undergoing Variational
sparsification. Regarding methods themselves, $\cplx$-VD and $\cplx$-ARD follow the same
declining compression-accuracy pattern, but for the same setting $C$ the latter provides
slightly less compression with marginally better accuracy.

@@

This study has direct implications for embedded deep learning applications
both in terms of lower storage requirements and higher throughput stemming from fewer
floating point multiplications due to sparsity, despite somewhat higher arithmetic
complexity of $\cplx$-valued networks.

@@

% the final camera-ready version can (and probably should) include acknowledgements
% Typically, this will include thanks to reviewers who gave useful comments, to colleagues
% who contributed to the ideas, and to funding agencies and corporate sponsors that provided
% financial support.
We would like to thank the anonymous reviewers, Evgenii Egorov, Ruslan Kostoev (ADASE)
and Danila Doroshin (Huawei) for their useful comments.
% We thank Skoltech CDISE HPC Zhores cluster staff for providing computational infrastructure.
The authors acknowledge the use of the Skoltech CDISE HPC cluster ``Zhores'' for obtaining
the results presented in this paper.
